---
title: "Live Data Stream Tutorial"
author: "K Duffy"
date: "5/13/2019"
output:
  html_document: default
  pdf_document: default
---
First let's load some packages:

```{r, silent=TRUE}
library(jsonlite)
library(phenocamapi)
library(plotly)
library(phenocamr)
library(dplyr)
```

As a refresher, there are two main ways to pull in PhenoCam data.  First, directly via the API:
```{r}
c      = jsonlite::fromJSON('https://phenocam.sr.unh.edu/api/cameras/?format=json&limit=2000')
c = c$results
c_m=c$sitemetadata
c$sitemetadata=NULL
cams_=cbind(c, c_m)
cams_[is.na(cams_)] = 'N'
cams_[, 2:4] <- sapply(cams_[, 2:4], as.numeric) #changing lat/lon/elev from string values into numeric
head(cams_)
```

And second, via the phenocamapi package:

```{r}
phenos=get_phenos()
head(phenos)
```

To familiarize yourself with the phenocam API, let's explore the structure:
[https://phenocam.sr.unh.edu/api/](https://phenocam.sr.unh.edu/api/)

Explore the options for filtering, file type and so forth.

***

Now, based on either direct API access or via the phenocamapi package, generate a dataframe of phenocam sites.  Select two phenocam sites from *different* plant functional types to explore (e.g. one grassland site and one evergreen needleleaf site)

```{r}
#example
GrassSites=cams_%>%
  filter(cams_$primary_veg_type=='GR')
head(GrassSites)
```

```{r}
FirstSite=GrassSites[5, ] #randomly chose the fifth site in the table
FirstSite
```

Chose your own sites and build out your code chunk here:
```{r}
print('build here')
```

[Koen Huffkens](https://khufkens.com/) developed the [phenocamr package](https://cran.r-project.org/web/packages/phenocamr/index.html), which streamlines access to quality controlled data.  We will now use this package to download and process site based data according to a standardized methodology.

A full description of the methodology is provided in Scientific Data: Tracking vegetation phenology across diverse North American biomes using PhenoCam imagery (Richardson et al. 2018).

```{r}
#uncomment if you need to install via devtools
#if(!require(devtools)){install.package(devtools)}
#devtools::install_github("khufkens/phenocamr")
library(phenocamr)
```


Use the dataframe of sites that you want to analyze to feed the phenocamr package.
Note: you can choose either a daily or 3 day product

```{r}
phenocamr::download_phenocam(
  frequency = 3,
  veg_type = FirstSite$primary_veg_type,
  roi_id = 1000,
  site = FirstSite$Sitename,
  phenophase = TRUE,
  out_dir = "."
  )
```

Now look in your working directory.  You have data!  Read it in:

```{r}
# load the time series data but replace the csv filename with whatever you downloaded
df <- read.table("butte_GR_1000_3day.csv", header = TRUE, sep = ",")

# read in the transition date file
td <- read.table("butte_GR_1000_3day_transition_dates.csv",
                 header = TRUE,
                 sep = ",")
```

Let's take a look at the data:

```{r}
p = plot_ly() %>%
  add_trace(
  data = df,
  x = ~ as.Date(date),
  y = ~ smooth_gcc_90,
  name = 'Smoothed GCC',
  showlegend = TRUE,
  type = 'scatter',
  mode = 'line'
) %>% add_markers(
  data=df,
  x ~ as.Date(date),
  y = ~gcc_90,
  name = 'GCC',
  type = 'scatter',
  color ='#07A4B5', 
  opacity=.5
)
p
```

What patterns do you notice?  How would we go about determining say the start of spring? (SOS)

###Threshold values

Let's subset the transition date (td) for each year when 25% of the greenness amplitude (of the 90^th) percentile is reached (threshold_25).

```{r}
# select the rising (spring dates) for 25% threshold of Gcc 90
spring <- td[td$direction == "rising" & td$gcc_value == "gcc_90",]

```

Now let's create a simple plot_ly line graph of the smooth Green Chromatic Coordinate (Gcc) and add points for transition dates:


```{r}
p = plot_ly() %>%
  add_trace(
  data = df,
  x = ~ as.Date(date),
  y = ~ smooth_gcc_90,
  name = 'PhenoCam GCC',
  showlegend = TRUE,
  type = 'scatter',
  mode = 'line'
) %>% add_markers(
  data= spring, 
  x = ~ as.Date(spring$transition_25, origin = "1970-01-01"),
  y = ~ spring$threshold_25,
  type = 'scatter',
  mode = 'marker',
  name = 'Spring Dates')
                
p
```

Now we can se the transition date for each year of interest and the annual patterns of the Gcc.

However, if you want more control over the parameters used during processing, you can run through the three default processing steps as implemented in download_phenocam() and set parameters manually.

Of particular interest is the option to specify your own threshold used in determining transition dates. 

Now, in your small groups:
What would be a reasonable threshold for peak greenness?  Or autumn onset?  Look at the ts dataset and phenocamr package and come up with a threshold.  Use the same code to plot it here:

```{r}
print('code here')
#some hint code
#what does 'rising' versus 'falling' denote?
#what threshold should you choose?
#td <- phenophases("butte_GR_1000_3day.csv",
#            internal = TRUE,
#            upper_thresh = 0.8)
fall <- td[td$direction == "falling" & td$gcc_value == "gcc_90",]
#Now generate a fall dataframe, what metrics should you use?
```

###Comparing phenology across vegetation types

Let's load in a function to make plotting smoother.  I'm dropped it here in the markdown so that you can edit it and re-run it as you see fit:
```{r}
gcc_plot = function(gcc, spring, fall){
  unix = "1970-01-01"

  p = plot_ly(
    data = gcc,
    x = ~ date,
    y = ~ gcc_90,
    showlegend = FALSE,
    type = 'scatter',
    mode = 'markers'
  ) %>%
    add_trace(
      y = ~ smooth_gcc_90,
      mode = "lines",
      line = list(width = 2, color = "rgb(120,120,120)"),
      name = "Gcc loess fit",
      showlegend = TRUE
    ) %>%
    # SOS spring
    # 10%
    add_trace(
      data = spring,
      x = ~ as.Date(transition_10),
      y = ~ threshold_10,
      mode = "markers",
      type = "scatter",
      marker = list(color = "#7FFF00", symbol = "circle"),
      name = "SOS (10%)",
      showlegend = TRUE
    ) %>%
    add_segments(x = ~ as.Date(transition_10_lower_ci),
                 xend = ~ as.Date(transition_10_upper_ci),
                 # y = ~ 0,
                 # yend = ~ 1,
                 y = ~ threshold_10,
                 yend = ~ threshold_10,
                 line = list(color = "#7FFF00"),
                 name = "SOS (10%) - CI"
    ) %>%
    # 25 %
    add_trace(
      x = ~ as.Date(transition_25),
      y = ~ threshold_25,
      mode = "markers",
      type = "scatter",
      marker = list(color = "#66CD00", symbol = "square"),
      showlegend = TRUE,
      name = "SOS (25%)"
    ) %>%
    add_segments(x = ~ as.Date(transition_25_lower_ci),
                 xend = ~ as.Date(transition_25_upper_ci),
                 y = ~ threshold_25,
                 yend = ~ threshold_25,
                 line = list(color = "#66CD00"),
                 name = "SOS (25%) - CI"
    ) %>%
    # 50 %
    add_trace(
      x = ~ as.Date(transition_50),
      y = ~ threshold_50,
      mode = "markers",
      type = "scatter",
      marker = list(color = "#458B00", symbol = "diamond"),
      showlegend = TRUE,
      name = "SOS (50%)"
    ) %>%
    add_segments(x = ~ as.Date(transition_50_lower_ci),
                 xend = ~ as.Date(transition_50_upper_ci),
                 y = ~ threshold_50,
                 yend = ~ threshold_50,
                 line = list(color = "#458B00"),
                 name = "SOS (50%) - CI"
    ) %>%
    
    # EOS fall
    # 50%
    add_trace(
      data = fall,
      x = ~ as.Date(transition_50),
      y = ~ threshold_50,
      mode = "markers",
      type = "scatter",
      marker = list(color = "#FFB90F", symbol = "diamond"),
      showlegend = TRUE,
      name = "EOS (50%)"
    ) %>%
    add_segments(x = ~ as.Date(transition_50_lower_ci),
                 xend = ~ as.Date(transition_50_upper_ci),
                 y = ~ threshold_50,
                 yend = ~ threshold_50,
                 line = list(color = "#FFB90F"),
                 name = "EOS (50%) - CI"
    ) %>%
    # 25 %
    add_trace(
      x = ~ as.Date(transition_25),
      y = ~ threshold_25,
      mode = "markers",
      type = "scatter",
      marker = list(color = "#CD950C", symbol = "square"),
      showlegend = TRUE,
      name = "EOS (25%)"
    ) %>%
    add_segments(x = ~ as.Date(transition_25_lower_ci),
                 xend = ~ as.Date(transition_25_upper_ci),
                 y = ~ threshold_25,
                 yend = ~ threshold_25,
                 line = list(color = "#CD950C"),
                 name = "EOS (25%) - CI"
    ) %>%
    # 10 %
    add_trace(
      x = ~ as.Date(transition_10),
      y = ~ threshold_10,
      mode = "markers",
      marker = list(color = "#8B6508", symbol = "circle"),
      showlegend = TRUE,
      name = "EOS (10%)"
    ) %>%
    add_segments(x = ~ as.Date(transition_10_lower_ci),
                 xend = ~ as.Date(transition_10_upper_ci),
                 y = ~ threshold_10,
                 yend = ~ threshold_10,
                 line = list(color = "#8B6508"),
                 name = "EOS (10%) - CI"
    )
  return (p)
}

```

```{r}

gcc_p = gcc_plot(df, spring, fall)
gcc_p
```

What is the difference in 25% greenness onset for your first site? #hint, look at the spring dataframe you just generated

```{r}
#some hints to get you started
d=spring$transition_25
d=as.Date(d)
d
```

```{r}
#more code hints
dates_split <- data.frame(date = d,
                 year = as.numeric(format(d, format = "%Y")),
                 month = as.numeric(format(d, format = "%m")),
                 day = as.numeric(format(d, format = "%d")))
```

***

Generate a plot of smoothed gcc and transition dates for your two sites and subplot them.  What do you notice?

```{r}
#some hint code for subplotting in plot_ly:
#p <- subplot(p1, p2, nrows=2)
#p
```

###Comparing phenology of the same vegetation cover but across climate space

As Dr. Richardson mentioned this morning in his introduction lecture, the same plant functional types (e.g. grasslands) can have very different phenologogical cycles.  Let's pick two phenocam grassland sites: one from a tropical climate (kamuela), and one from an arid climate (konza) here in Arizona:

```{r}
GrassSites=GrassSites[c(27,32),]
```

Now use the code you've generated above to pull in data from those sites:

```{r}
print('code here')
```

Now let's create a subplot of your grasslands to compare phenology, some hint code below:

```{r}
#some hint code for subplotting in plot_ly:
#p <- subplot(p1, p2, nrows=2)
#p
```

Once you have a subplot of grassland phenology across 2 climates answer the following questions here in the markdown:
1. What seasonal patterns do you see?
2. Do you think you set your thresholds correctly for transition dates/phenophases?  How might that very as a function of climate?
3. What are the challenges of forecasting or modeling tropical versus arid grasslands?  How could the material that you've learned in this course so far aid you?