---
title: "Live Data Stream Tutorial"
author: "K Duffy"
date: "5/13/2019"
output:
  html_document: default
  pdf_document: default
---
First let's load some packages:

```{r, silent=TRUE}
library(jsonlite)
library(phenocamapi)
library(plotly)
library(phenocamr)
library(dplyr)
```

As a refresher, there are two main ways to pull in PhenoCam data.  First, directly via the API:
```{r}
c      = jsonlite::fromJSON('https://phenocam.sr.unh.edu/api/cameras/?format=json&limit=2000')
c = c$results
c_m=c$sitemetadata
c$sitemetadata=NULL
cams_=cbind(c, c_m)
cams_[is.na(cams_)] = 'N'
cams_[, 2:4] <- sapply(cams_[, 2:4], as.numeric) #changing lat/lon/elev from string values into numeric
head(cams_)
```

And second, via the phenocamapi package:

```{r}
phenos=get_phenos()
head(phenos)
```

To familiarize yourself with the phenocam API, let's explore the structure:
[https://phenocam.sr.unh.edu/api/](https://phenocam.sr.unh.edu/api/)

Explore the options for filtering, file type and so forth.

***

Now, based on either direct API access or via the phenocamapi package, generate a dataframe of phenocam sites.  Select two phenocam sites from *different* plant functional types to explore (e.g. one grassland site and one evergreen needleleaf site)

```{r}
#example
GrassSites=cams_%>%
  filter(cams_$primary_veg_type=='GR')
head(GrassSites)
```

```{r}
FirstSite=GrassSites[5, ] #randomly chose the fifth site in the table
FirstSite
```

Chose your own sites and build out your code chunk here:
```{r}
print('build here')
```

[Koen Huffkens](https://khufkens.com/) developed the [phenocamr package](https://cran.r-project.org/web/packages/phenocamr/index.html), which streamlines access to quality controlled data.  We will now use this package to download and process site based data and process it according to a standardized methodology.

A full description of the methodology is provided in Scientific Data: Tracking vegetation phenology across diverse North American biomes using PhenoCam imagery (Richardson et al. 2018).

```{r}
#uncomment if you need to install via devtools
#if(!require(devtools)){install.package(devtools)}
#devtools::install_github("khufkens/phenocamr")
library(phenocamr)
```


Use the dataframe of sites that you want to analyze to feed the phenocamr package.
Note: you can choose either a daily or 3 day product

```{r}
phenocamr::download_phenocam(
  frequency = 3,
  veg_type = FirstSite$primary_veg_type,
  roi_id = 1000,
  site = FirstSite$Sitename,
  phenophase = TRUE,
  out_dir = "."
  )
```

Now look in your working directory.  You have data!  Read it in:

```{r}
# load the time series data but replace the csv filename with whatever you downloaded
df <- read.table("butte_GR_1000_3day.csv", header = TRUE, sep = ",")

# read in the transition date file
td <- read.table("butte_GR_1000_3day_transition_dates.csv",
                 header = TRUE,
                 sep = ",")
```

```{r}
# select the rising (spring dates) for 25% threshold of Gcc 90
spring <- td[td$direction == "rising" & td$gcc_value == "gcc_90",]

```


Now let's create a simple plot_ly line graph of the smooth Green Chromatic Coordinate (Gcc) and add points for transition dates:

Generate a phenocam plotting function to streamline:
```{r}
p = plot_ly() %>%
  add_trace(
  data = df,
  x = ~ as.Date(date),
  y = ~ smooth_gcc_90,
  name = 'PhenoCam GCC',
  showlegend = TRUE,
  type = 'scatter',
  mode = 'line'
) %>% add_markers(
  data= spring, 
  x = ~ as.Date(spring$transition_25, origin = "1970-01-01"),
  y = ~ spring$threshold_25,
  type = 'scatter',
  mode = 'marker',
  name = 'Spring Dates')
                
p
```

Now, in your small groups:
What would be a reasonable threshold for peak greenness?  Or autumn onset?  Look at the ts dataset and phenocamr package and come up with a threshold.  Use the same code to plot it here:

```{r}
print('code here')
#some hint code
#what does 'rising' versus 'falling' denote?
#what threshold should you choose?
#td <- phenophases("butte_GR_1000_3day.csv",
#            internal = TRUE,
#            upper_thresh = 0.8)
fall <- td[td$direction == "falling" & td$gcc_value == "gcc_90",]
#Now generate a fall dataframe, what metrics should you use?
```

###Comparing phenology across vegetation types


```{r}
gcc_plot = function(gcc, spring, fall){
  unix = "1970-01-01"

  p = plot_ly(
    data = gcc,
    x = ~ date,
    y = ~ gcc_90,
    showlegend = FALSE,
    type = 'scatter',
    mode = 'markers'
  ) %>%
    add_trace(
      y = ~ smooth_gcc_90,
      mode = "lines",
      line = list(width = 2, color = "rgb(120,120,120)"),
      name = "Gcc loess fit",
      showlegend = TRUE
    ) %>%
    # SOS spring
    # 10%
    add_trace(
      data = spring,
      x = ~ as.Date(transition_10),
      y = ~ threshold_10,
      mode = "markers",
      type = "scatter",
      marker = list(color = "#7FFF00", symbol = "circle"),
      name = "SOS (10%)",
      showlegend = TRUE
    ) %>%
    add_segments(x = ~ as.Date(transition_10_lower_ci),
                 xend = ~ as.Date(transition_10_upper_ci),
                 # y = ~ 0,
                 # yend = ~ 1,
                 y = ~ threshold_10,
                 yend = ~ threshold_10,
                 line = list(color = "#7FFF00"),
                 name = "SOS (10%) - CI"
    ) %>%
    # 25 %
    add_trace(
      x = ~ as.Date(transition_25),
      y = ~ threshold_25,
      mode = "markers",
      type = "scatter",
      marker = list(color = "#66CD00", symbol = "square"),
      showlegend = TRUE,
      name = "SOS (25%)"
    ) %>%
    add_segments(x = ~ as.Date(transition_25_lower_ci),
                 xend = ~ as.Date(transition_25_upper_ci),
                 y = ~ threshold_25,
                 yend = ~ threshold_25,
                 line = list(color = "#66CD00"),
                 name = "SOS (25%) - CI"
    ) %>%
    # 50 %
    add_trace(
      x = ~ as.Date(transition_50),
      y = ~ threshold_50,
      mode = "markers",
      type = "scatter",
      marker = list(color = "#458B00", symbol = "diamond"),
      showlegend = TRUE,
      name = "SOS (50%)"
    ) %>%
    add_segments(x = ~ as.Date(transition_50_lower_ci),
                 xend = ~ as.Date(transition_50_upper_ci),
                 y = ~ threshold_50,
                 yend = ~ threshold_50,
                 line = list(color = "#458B00"),
                 name = "SOS (50%) - CI"
    ) %>%
    
    # EOS fall
    # 50%
    add_trace(
      data = fall,
      x = ~ as.Date(transition_50),
      y = ~ threshold_50,
      mode = "markers",
      type = "scatter",
      marker = list(color = "#FFB90F", symbol = "diamond"),
      showlegend = TRUE,
      name = "EOS (50%)"
    ) %>%
    add_segments(x = ~ as.Date(transition_50_lower_ci),
                 xend = ~ as.Date(transition_50_upper_ci),
                 y = ~ threshold_50,
                 yend = ~ threshold_50,
                 line = list(color = "#FFB90F"),
                 name = "EOS (50%) - CI"
    ) %>%
    # 25 %
    add_trace(
      x = ~ as.Date(transition_25),
      y = ~ threshold_25,
      mode = "markers",
      type = "scatter",
      marker = list(color = "#CD950C", symbol = "square"),
      showlegend = TRUE,
      name = "EOS (25%)"
    ) %>%
    add_segments(x = ~ as.Date(transition_25_lower_ci),
                 xend = ~ as.Date(transition_25_upper_ci),
                 y = ~ threshold_25,
                 yend = ~ threshold_25,
                 line = list(color = "#CD950C"),
                 name = "EOS (25%) - CI"
    ) %>%
    # 10 %
    add_trace(
      x = ~ as.Date(transition_10),
      y = ~ threshold_10,
      mode = "markers",
      marker = list(color = "#8B6508", symbol = "circle"),
      showlegend = TRUE,
      name = "EOS (10%)"
    ) %>%
    add_segments(x = ~ as.Date(transition_10_lower_ci),
                 xend = ~ as.Date(transition_10_upper_ci),
                 y = ~ threshold_10,
                 yend = ~ threshold_10,
                 line = list(color = "#8B6508"),
                 name = "EOS (10%) - CI"
    )
  return (p)
}

```

```{r}

gcc_p = gcc_plot(df, spring, fall)
gcc_p
```
